Management induced changes of soil organic carbon on global croplands

### 2022-07-20 ResultNotebooks/results_SOCscenario.Rmd
update to version before bug check
update to final version - 2. resubmit

theme_set(theme_bw())
legend_size <- 10
scen_cols   <- c("#648FFF", "#785EF0", "#DC267F")

@@

startY <- 1965

initSOCemis <- 
  ggplot() + 
  geom_path(data = as.data.frame(-glo_SOCDebt[, seq(startY, 2010, 1), ]),
            aes(x = Year, y = Value, group = interaction(Data1, Data2), 
                linetype = Data1, color = Data2)) +
  labs(x = "Year",
        y = expression(paste("Global \u0394SOC in GtC")),
        color    = "Initialization",
        linetype = "Management") +
  scale_x_discrete(breaks = seq(startY, 2010, 5)) +
  scale_y_continuous(limits = limits, breaks = breaks) +
  scale_color_manual(values = c(scen_cols), 
                     labels = unique(gsub("(.*)\\.(.*)", "\\2", 
                                          names(init_scen)))) +
  scale_linetype_manual(values = c(1, 2), 
                        labels = unique(gsub("(.*)\\.(.*)", "\\1", 
                                             names(init_scen)))) +
  theme(rect = element_blank(),
        axis.title.y = element_text(size = legend_size, 
                                    margin = margin(t = 0, r = 10, b = 0, l = 0, 
                                                    unit = "pt")),
        axis.title.x = element_text(size = legend_size, 
                                    margin = margin(t = 10, r = 0, b = 0, l = 0, 
                                                    unit = "pt")),
        axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8, 
                                   size = legend_size),
        legend.position = "right", legend.direction = "vertical")

@@

ggsave(paste0(outputdir, "Images/scenario_init.png"), 
       plot = initSOCemis, width = 12, height = 8, dpi = 300, units = "cm")
```

>>>>

theme_set(theme_bw())
legend_size <- 12
scen_cols   <- c("#648FFF", "#785EF0", "#DC267F")

@@

startY <- 1975

initSOCemis <- 
  ggplot() + 
  geom_path(data = as.data.frame(-glo_SOCDebt[, seq(startY, 2010, 1), ]),
            aes(x = Year, y = Value, group = interaction(Data1, Data2), 
                linetype = Data1, color = Data2)) +
  labs(x = "Year",
        y = expression(paste("Global \u0394SOC in GtC")),
        color    = "Initialization",
        linetype = "Management") +
  scale_x_discrete(breaks = seq(startY, 2010, 5)) +
  scale_y_continuous(limits = limits, breaks = breaks) +
  scale_color_manual(values = c(scen_cols), 
                     labels = unique(gsub("(.*)\\.(.*)", "\\2", 
                                          names(init_scen)))) +
  scale_linetype_manual(values = c(1, 2), 
                        labels = unique(gsub("(.*)\\.(.*)", "\\1", 
                                             names(init_scen)))) +
  theme(rect = element_blank(),
        axis.title.y = element_text(size = legend_size, 
                                    margin = margin(t = 0, r = 10, b = 0, l = 0, 
                                                    unit = "pt")),
        axis.title.x = element_text(size = legend_size, 
                                    margin = margin(t = 10, r = 0, b = 0, l = 0, 
                                                    unit = "pt")),
        axis.text.x  = element_text(angle = 60, hjust = 0.8, vjust = 0.8, 
                                   size = legend_size),
        axis.text.y  = element_text(size = legend_size),
        legend.title = element_text(size = legend_size),
        legend.text  = element_text(size = legend_size),
        legend.position = "right", legend.direction = "vertical")

@@

ggsave(paste0(outputdir, "Images/scenario_init.png"), 
       plot = initSOCemis, width = 15, height = 10, dpi = 300, units = "cm")
```

### 2022-07-20 SOCbudgetPaper/SOCbudget.Rmd
update to final version - 2. resubmit

  Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work. Additional thanks to the rticles contributors for providing a R Markdown template. The authors thank for the data provided by FAOSTAT and LUH2v2. The work of KK was funded by the DFG Priority Program “Climate Engineering: Risks, Challenges, Opportunities?” (SPP 1689) and specifically the CEMICS2 project (grant no. ED78/3-2). The research leading to these results has received funding for BLB from the European Union’s Horizon 2020 research and innovation program under grant agreement no. 776479 (COACCH) and  no. 821010 (CASCADES). The work of SR, JS and IW was also supported by CLIMASTEPPE (01DJ8012), EXIMO (01LP1903D) and FOCUS (031B0787B) funded by the German Federal Ministry of Education and Research (BMBF). The input of PS, MK and MD contributes to the Soils-R-GGREAT project (NE/P019455/1) and CIRCASA (EU H2020; grant agreement no. 774378).
appendix: |

  \section{Figures and tables in appendices}
  \subsection{Methods}
      ```{=latex}
  \begin{figure}[h]
    \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones} 
    \caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \citep{eggleston_ipcc_2006} and has been reimplented by \citet{carre_background_2010}, which is the source of this data. Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange.}
    \label{fig:CLIMzone}
  \end{figure}
  ```

  ```{=latex}
  \begin{sidewaystable*}[htbp]
  \caption{Parameterization of harvested organs and their corresponding residues parts as well as allometric coefficients: This table is mainly based on \citet{bodirsky_n2o_2012} together with simple carbon to dry matter assumptions. Allometric coefficients are used as descriped in \citet{eggleston_ipcc_2006} with HI\textsuperscript{prod} being slope\textsubscript{(T)}, HI\textsuperscript{area} intercept\textsubscript{(T)} and RS R\textsubscript{BG-BIO}.
  }
  \begin{tabular}{lllllllllllll}
  \tophline

@@

  <!-- closs in AWMS - Table {#append:TableclossAWMS} -->

@@

    ```{=latex}
  \begin{figure}[H]
    \includegraphics[width=12cm]{../ResultNotebooks/Output/Images/scenario_init.png} 
    \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting in 1901 with steady-state SOC under vegetation for all land-use types without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use (Initial-spinup1510). On the other hand, assuming all SOC to be in land-use specific steady-state already in 1901 (Initial-eq) leads to the opposite effect of an already higher $\Delta SOC$ in 1975 and a less steeper increase till 2010. However, the difference between the dynamic historical management assumption (histManagement) compared to constant management assumption from 1975 on (constManagement) are of a similar size for all initialization choices.}
    \label{fig:SOCinit}
  \end{figure}
  ```

    ```{=latex}
  \begin{figure}[H]
    \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure.png} 
    \caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations. In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010. PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model. We use the climate zone specification of the IPCC \citep{eggleston_ipcc_2006}.}
    \label{fig:SOCreg}
  \end{figure}
  ```

    ```{=latex}
  \begin{figure}[H]
    \includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scatterComparePointData_2010.png} 
    \caption{Correlation between modeled and measured $SOC$ stocks. Given the wide span between minimun and maximum measured $SOC$ stocks within in a given cell, we correlated median values with our modeled results. Both cropland ($R2=0.09$) and areas with natural vegetation ($R2=0.08$) tend to be lower in our results than in the point measurements.}
    \label{fig:SOCpoint}
  \end{figure}
   ```

     ```{=latex}
  \begin{figure}[H]
    \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/scatterCompareLPJmL42010_pnv.png}
  \caption{Correlation between modeled SOC stocks of LPJmL4 and this study for an hypothetical potential natural state (PNV) for the year 2010. The grey lines indicate the 1:1 line.}
    \label{fig:SOCLPJmL4}
  \end{figure}
   ```

>>>>

  Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work. Additional thanks to the rticles contributors for providing a R Markdown template. The authors thank for the data provided by FAOSTAT and LUH2v2. The work of KK was funded by the DFG Priority Program “Climate Engineering: Risks, Challenges, Opportunities?” (SPP 1689) and specifically the CEMICS2 project (grant no. ED78/3-2). The research leading to these results has received funding for BLB from the European Union’s Horizon 2020 research and innovation program under grant agreement no. 776479 (COACCH) and  no. 821010 (CASCADES). The work of SR, JS and IW was also supported by CLIMASTEPPE (01DJ8012), EXIMO (01LP1903D) and FOCUS (031B0787B) funded by the German Federal Ministry of Education and Research (BMBF). The input of PS, MK and MD contributes to the Soils-R-GGREAT project (NE/P019455/1) and CIRCASA (EU H2020; grant agreement no. 774378).
appendix: |
  \section{Figures and tables in appendices}

  \subsection{Methods}

    ```{=latex}
    \begin{figure}[h]
    \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones.png} 
    \caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \citep{eggleston_ipcc_2006} and has been reimplented by \citet{carre_background_2010}, which is the source of this data. Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange.}
    \label{fig:CLIMzone}
    \end{figure}
    ```

  ```{=latex}
  \begin{sidewaystable*}[htbp]
  \caption{Parameterization of harvested organs and their corresponding residues parts as well as allometric coefficients: This table is mainly based on \citet{bodirsky_n2o_2012} together with simple carbon to dry matter assumptions. Allometric coefficients are used as descriped in \citet{eggleston_ipcc_2006} with HI\textsuperscript{prod} being slope\textsubscript{(T)}, HI\textsuperscript{area} intercept\textsubscript{(T)} and RS R\textsubscript{BG-BIO}.
  }
  \begin{tabular}{lllllllllllll}
  \tophline

@@

@@

    ```{=latex}
    \begin{figure}[h]
    \includegraphics[width=12cm]{../ResultNotebooks/Output/Images/scenario_init.png} 
    \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting in 1901 with steady-state SOC under vegetation for all land-use types without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use (Initial-spinup1510). On the other hand, assuming all SOC to be in land-use specific steady-state already in 1901 (Initial-eq) leads to the opposite effect of an already higher $\Delta SOC$ in 1975 and a less steeper increase till 2010. However, the difference between the dynamic historical management assumption (histManagement) compared to constant management assumption from 1975 on (constManagement) are of a similar size for all initialization choices.}
    \label{fig:SOCinit}
    \end{figure}
    ```

    ```{=latex}
    \begin{figure}[h]
    \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure.png} 
    \caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations. In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010. PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model. We use the climate zone specification of the IPCC \citep{eggleston_ipcc_2006}.}
    \label{fig:SOCreg}
    \end{figure}
    ```

    ```{=latex}
    \begin{figure}[h]
    \includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scatterComparePointData_2010.png} 
    \caption{Correlation between modeled and measured $SOC$ stocks. Given the wide span between minimun and maximum measured $SOC$ stocks within in a given cell, we correlated median values with our modeled results. Both cropland ($R2=0.09$) and areas with natural vegetation ($R2=0.08$) tend to be lower in our results than in the point measurements.}
    \label{fig:SOCpoint}
    \end{figure}
    ```

    ```{=latex}
     \begin{figure}[h]
    \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/scatterCompareLPJmL42010_pnv.png}
  \caption{Correlation between modeled SOC stocks of LPJmL4 and this study for an hypothetical potential natural state (PNV) for the year 2010. The grey lines indicate the 1:1 line.}
    \label{fig:SOCLPJmL4}
    \end{figure}
   ```

### 2022-07-20 SOCbudgetPaper/SOCbudget.tex
update to final version - 2. resubmit

According to the IPCC guidelines \(SOC\) changes can be expressed as the difference of two consecutive years \citep[see Eq. 5.0A in][]{ogle_cropland_in_ipcc_2019}. This, however, will also include naturally occurring changes due to climatic variation over time. For our study, we define the absolute and relative \(SOC\) changes in relation to a potential natural state \(SOC^{\mathrm{pnv}}\) under the same climatic conditions in grid cell \(i\) at time \(t\), that is based on the natural vegetation \(SOC\) calculations as defined above without accounting for land conversion from cropland at any time. The absolute changes \(\Delta SOC\) and relative changes \(F^{\mathrm{SCF}}\) are thus given by

@@

From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in Bodirsky et al. \citeyearpar{bodirsky_mrcommons_2020}). Residue feed demand for five different livestock groups is based on country-specific feed baskets \citep[see][]{weindl_livestock_2017}, that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account. We estimate a material-use share for the straw-residue group of 5\% and a fuel-share of 10\% for all used residue groups in low-income countries. For high-income countries, no withdrawal for material or fuel use is assumed, and use shares of middle-income countries are linearly interpolated based on per-capita income, following the same rationale as for the share of burnt residues described above. The remaining \(AGR\) as well as all \(BGR\) are expected to be left on the field. We limit high residue return rates to at most \(10\,\unit{tC\,ha}^{-1}\) in order to correct for outliers.

@@

Manure production and management is based on a revised methodology of Bodirsky et al. \citeyearpar{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modeling. Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted. Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see Calvo Buendia et al. \citeyearpar{ogle_cropland_in_ipcc_2019}.
Depending on the feed system we assume manure to be handled in four different ways:
All manure originated from pasture feed intake is excreted directly on pastures and rangelands (pasture grazing), deducting manure collected as fuel.
Whereas for low-income countries, we adopt a share of 25\% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
For all other feed items, we assume the manure to be stored in animal waste management systems associated with livestock housing.
To estimate the carbon actually returned to the soil, we account for carbon losses during storage, where return shares depend on different animal waste management and grazing systems. Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields. For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in Bodirsky et al. \citeyearpar{bodirsky_mrcommons_2020} for more details) depending on the different systems and the associated nitrogen loss rates as specified in Bodirsky et al. \citeyearpar{bodirsky_n2o_2012}. We limit high application shares at \(10\,\unit{tC\,ha}^{-1}\) to correct for outliers, that can occur due to inconsistencies between FAO production and 0.5 degree land-use data.

@@

In Fig. \ref{fig:SOCmaps}(b) we provide a world map of SOC stocks estimates for the first \(30\,\unit{cm}\) on cropland considering historical management data for the year 2010. Values range between over \(100\,\unit{t\,ha^{-1}}\) in northern temperate cropland to less than \(5\,\unit{t\,ha^{-1}}\) for arid and semiarid cropland.
Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under PNV in two complementary ways:
1. Absolute SOC changes \(\Delta SOC\) (Fig. \ref{fig:SOCmaps}(c)) indicate areas with high importance for the global \(SOC\) loss. They can be driven by large relative changes (e.g.~in Central Africa) or by a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.~North-East Asia).
2. Relative SOC changes measured as stock changes factors \(F^{SCF}\) (Fig. \ref{fig:SOCmaps}(d)) are a helpful metric to analyze the impact of human cropping activities. They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome by improved agricultural practices. Large parts of tropical cropland seem to suffer from strongly reduced relative stocks, indicating SOC degradation. Conversely, irrigated cropland at the border to dry, unsuitable areas worldwide shows a strong relative increase in SOC stocks.

@@

Using the constManagement results, that only include land-use change (LUC) related changes of the SOC debt between 1975 and 2010, we are able to subtract the LUC effect from the overall SOC debt change within the histManagement results. The remaining effect can be attributed to the changing agricultural management (MAN) as other drivers such as climatic effects have been already canceled out by taking the difference to a PNV reference state while calculating \(\Delta SOC\).
The increasing SOC debt on global cropland are primarily caused by LUC (red areas in Fig. \ref{fig:SOCdebt}(b2)). Deteriorated management also contributed to increasing SOC debt in parts of Sub-saharan Africa and Central Asia. In contrast, agricultural management has led to an decrease in SOC debt in the USA, Europe, as well as in parts of China and India (blue areas in Fig. \ref{fig:SOCdebt}(b3)), which is not visible in the total \(\Delta SOC\) change as LUC was happening at the same time.

@@

To evaluate our modeled SOC stocks and stock changes under agricultural management, we compare our results to the default IPCC stock change factors \(F^{\mathrm{SCF}}\) of 2006 \citep{lasco_cropland_in_ipcc_2006} and their refinements in 2019 \citep{ogle_cropland_in_ipcc_2019}. Both estimates are based on measurement data for cropland (see Fig. \ref{fig:SCFglo}). To allow for comparison, we aggregate our stock change factors weighted by grid-level cropland area to derive median factors for the four IPCC climate zones (Fig. \ref{fig:CLIMzone}). Note that IPCC Tier 1 factors are derived under the assumption that there is a linear change between steady states over 20 years, whereas our aggregated factors just reflect the relative change compared to a given potential natural vegetation reference stock without specifically tracking the time scale of the stock change.

\begin{figure}[h]
\includegraphics[width=15cm]{../ResultNotebooks/Output/Images/SCF_comparison} \caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \citep{ogle_cropland_in_ipcc_2019}. IPCC factors vary a bit with different types of management (low, medium, high). The results of the study span over a wide range -- indicated by the span between the 5th and 95th percentile -- but not overlapping with the IPCCs 2019 factors for tropical moist. The median for most regions is considerably lower than the IPCC factors.}\label{fig:SCFglo}
\end{figure}

@@

The global estimates of the total SOC stock of the upper 30 cm from this study are in the middle of the wide range of other modeled or observation-based estimates. Regional results (Fig. \ref{fig:SOCreg}) show that our estimates are well within the range of other estimates for most regions, but at the lower end for tropical moist and tropical wet areas. SoilGrids \citep{hengl_soilgrids250m_2017} especially stands out with its high estimate, since they include the litter horizon on top of the soil, that might dominate especially polar and boreal soils. SoilGrids 2.0 \citep{poggio_soilgrids_2021} however, excludes litter C and thus marks the lower end for particularly northern regions. For the same reason it is also more comparable to our results, which do not account for litter C as well.

@@

Estimates of SOC stocks under natural vegetation influence our modeled results for cropland, which has been converted from natural vegetation at some point in time. As the Tier 2 modeling approach \citep{ogle_cropland_in_ipcc_2019} is not specifically parameterized for natural vegetation it is important to evaluate its suitability to produce reasonable results in that domain as well at least comparable to other modeling approaches.
We therefore also compare our modeled results for SOC under natural vegetation (derived using litterfall of LPJmL4) against estimates of SOC by LPJmL4 for a PNV simulation (see Fig.\ref{fig:SOCLPJmL4}). Both models are driven by the same climate conditions and the same natural litterfall and just differ in the representation of SOC and litter dynamics. With our focus on cropland SOC dynamics, we compare only cells with more than \(1000\,\unit{ha}\) of cropland (capturing 99.9\% of global cropland area).
Spatial correlations of PNV SOC stock values are high (global \(R2=0.81\)), especially for dry climate zones (Fig. \ref{fig:SOCLPJmL4}). For temperate and tropical moist areas estimates of this study tend to be a bit lower compared to LPJmL4 results.

@@

The Tier 2 approach \citep{ogle_cropland_in_ipcc_2019} used here is explicitly designed for agricultural soils, whereas we apply it also to soils under PNV. This is necessary in order to represent SOC losses under land-use change in a dynamic way, as this is an important driver of SOC dynamics. The comparison of simulated PNV data with LPJmL4 shows the model's capability in reproducing PNV SOC stocks (Fig. \ref{fig:SOCLPJmL4}).
Concurrently, the point-data comparison (see Fig. \ref{fig:SOCpoint}) shows low correlation for PNV, however also for cropland sites. This might also point to the fact that SOC stocks can vary at field and local scale considerably and thus a very high number of point data is needed to derive statistical properties, that could improve the point-to-grid-cell comparison \citep[see][]{rammig_generic_2018}.f

@@

The initialization of SOC stocks, however, is important for the size of the SOC debt and its change over time, since the presence and size of legacy fluxes effect these values strongly (see Fig. \ref{fig:SOCinit}). According to our sensitivity analysis the SOC debt depending on the initialization choice has been between \(33.3\,\unit{GtC}\) and \(50.7\,\unit{GtC}\), with our best guess at \(39.6\,\unit{GtC}\). Concurrently, our results indicate, that the impact of the dynamic agricultural management is robust to the initialization of SOC stocks.

>>>>

According to the IPCC guidelines \(SOC\) changes can be expressed as the difference of two consecutive years \citep[see Eq. 5.0A in][]{ogle_cropland_in_ipcc_2019}. This, however, will also include naturally occurring changes due to climatic variation over time. For our study, we define the absolute and relative \(SOC\) changes in relation to a potential natural state \(SOC^{\mathrm{pnv}}\) under the same climatic conditions in grid cell \(i\) at time \(t\) that is based on the natural vegetation \(SOC\) calculations as defined above without accounting for land conversion from cropland at any time. The absolute changes \(\Delta SOC\) and relative changes \(F^{\mathrm{SCF}}\) are thus given by

@@

From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in Bodirsky et al. \citeyearpar{bodirsky_mrcommons_2020}). Residue feed demand for five different livestock groups is based on country-specific feed baskets \citep[see][]{weindl_livestock_2017} that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account. We estimate a material-use share for the straw-residue group of 5\% and a fuel-share of 10\% for all used residue groups in low-income countries. For high-income countries, no withdrawal for material or fuel use is assumed, and use shares of middle-income countries are linearly interpolated based on per-capita income, following the same rationale as for the share of burnt residues described above. The remaining \(AGR\) as well as all \(BGR\) are expected to be left on the field. We limit high residue return rates to at most \(10\,\unit{tC\,ha}^{-1}\) in order to correct for outliers.

@@

Manure production and management is based on a revised methodology of Bodirsky et al. \citeyearpar{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modeling. Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted. Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see Calvo Buendia et al. \citeyearpar{ogle_cropland_in_ipcc_2019}.
Depending on the feed system we assume manure to be handled in four different ways:
All manure originated from pasture feed intake is excreted directly on pastures and rangelands (pasture grazing), deducting manure collected as fuel.
Whereas for low-income countries, we adopt a share of 25\% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
For all other feed items, we assume the manure to be stored in animal waste management systems associated with livestock housing.
To estimate the carbon actually returned to the soil, we account for carbon losses during storage, where return shares depend on different animal waste management and grazing systems. Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields. For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in Bodirsky et al. \citeyearpar{bodirsky_mrcommons_2020} for more details) depending on the different systems and the associated nitrogen loss rates as specified in Bodirsky et al. \citeyearpar{bodirsky_n2o_2012}. We limit high application shares at \(10\,\unit{tC\,ha}^{-1}\) to correct for outliers that can occur due to inconsistencies between FAO production and 0.5 degree land-use data.

@@

In Fig. \ref{fig:SOCmaps}(b) we provide a world map of SOC stocks estimates for the first \(30\,\unit{cm}\) on cropland considering historical management data for the year 2010. Values range between over \(100\,\unit{t\,ha^{-1}}\) in northern temperate cropland to less than \(5\,\unit{t\,ha^{-1}}\) for arid and semiarid cropland.
Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under PNV in two complementary ways:
1. Absolute SOC changes \(\Delta SOC\) (Fig. \ref{fig:SOCmaps}(c)) indicate areas with high importance for the global \(SOC\) loss. They can be driven by large relative changes (e.g.~in Central Africa) or by a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.~North-East Asia).
2. Relative SOC changes measured as stock changes factors \(F^{SCF}\) (Fig. \ref{fig:SOCmaps}(d)) are a helpful metric to analyze the impact of human cropping activities. They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation that may hold potential to be overcome by improved agricultural practices. Large parts of tropical cropland seem to suffer from strongly reduced relative stocks, indicating SOC degradation. Conversely, irrigated cropland at the border to dry, unsuitable areas worldwide shows a strong relative increase in SOC stocks.

@@

Using the constManagement results that only include land-use change (LUC) related changes of the SOC debt between 1975 and 2010, we are able to subtract the LUC effect from the overall SOC debt change within the histManagement results. The remaining effect can be attributed to the changing agricultural management (MAN) as other drivers such as climatic effects have been already canceled out by taking the difference to a PNV reference state when calculating \(\Delta SOC\).
The increasing SOC debt on global cropland are primarily caused by LUC (red areas in Fig. \ref{fig:SOCdebt}(b2)). Deteriorated management also contributed to increasing SOC debt in parts of Sub-saharan Africa and Central Asia. In contrast, agricultural management has led to an decrease in SOC debt in the USA, Europe, as well as in parts of China and India (blue areas in Fig. \ref{fig:SOCdebt}(b3)), which is not visible in the total \(\Delta SOC\) change as LUC was happening at the same time.

@@

To evaluate our modeled SOC stocks and stock changes under agricultural management, we compare our results to the default IPCC stock change factors \(F^{\mathrm{SCF}}\) of 2006 \citep{lasco_cropland_in_ipcc_2006} and their refinements in 2019 \citep{ogle_cropland_in_ipcc_2019}. Both estimates are based on measurement data for cropland (see Fig. \ref{fig:SCFglo}). To allow for comparison, we aggregate our stock change factors weighted by grid-level cropland area to derive median factors for the four IPCC climate zones (Fig. \ref{fig:CLIMzone}). Note that IPCC Tier 1 factors are derived under the assumption that there is a linear change between steady states over 20 years, whereas our aggregated factors just reflect the relative change compared to a given potential natural vegetation reference stock without specifically tracking age of the cropland.

\begin{figure}[h]
\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/SCF_comparison} \caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \citep{ogle_cropland_in_ipcc_2019}. IPCC factors vary a bit with different types of management (low, medium, high). The results of the study span over a wide range -- indicated by the span between the 5th and 95th percentile -- but not overlapping with the IPCCs 2019 factors for tropical moist. The median for most regions is considerably lower than the IPCC factors.}\label{fig:SCFglo}
\end{figure}

@@

The global estimates of the total SOC stock of the upper 30 cm from this study are in the middle of the wide range of other modeled or observation-based estimates. Regional results (Fig. \ref{fig:SOCreg}) show that our estimates are well within the range of other estimates for most regions, but at the lower end for tropical moist and tropical wet areas. SoilGrids \citep{hengl_soilgrids250m_2017} especially stands out with its high estimate, since they include the litter horizon on top of the soil that might dominate especially polar and boreal soils. SoilGrids 2.0 \citep{poggio_soilgrids_2021} however, excludes litter C and thus marks the lower end for particularly northern regions. For the same reason it is also more comparable to our results, which do not account for litter C as well.

@@

Estimates of SOC stocks under natural vegetation influence our modeled results for cropland, which has been converted from natural vegetation at some point in time. As the Tier 2 modeling approach \citep{ogle_cropland_in_ipcc_2019} is not specifically parameterized for natural vegetation it is important to evaluate its suitability to produce reasonable results in that domain as well at least comparable to other modeling approaches.
We therefore also compare our modeled results for SOC under natural vegetation (derived using litterfall of LPJmL4) against estimates of SOC by LPJmL4 for a PNV simulation (see Fig. \ref{fig:SOCLPJmL4}). Both models are driven by the same climate conditions and the same natural litterfall and just differ in the representation of SOC and litter dynamics. With our focus on cropland SOC dynamics, we compare only cells with more than \(1000\,\unit{ha}\) of cropland (capturing 99.9\% of global cropland area).
Spatial correlations of PNV SOC stock values are high (global \(R2=0.81\)), especially for dry climate zones (Fig. \ref{fig:SOCLPJmL4}). For temperate and tropical moist areas estimates of this study tend to be a bit lower compared to LPJmL4 results.

@@

The Tier 2 approach \citep{ogle_cropland_in_ipcc_2019} used here is explicitly designed for agricultural soils, whereas we apply it also to soils under PNV. This is necessary in order to represent SOC losses under land-use change in a dynamic way, as this is an important driver of SOC dynamics. The comparison of simulated PNV data with LPJmL4 shows the model's capability in reproducing PNV SOC stocks (Fig. \ref{fig:SOCLPJmL4}).
Concurrently, the point-data comparison (see Fig. \ref{fig:SOCpoint}) shows low correlation for PNV, however also for cropland sites. This might also point to the fact that SOC stocks can vary at field and local scale considerably and thus a very high number of point data is needed to derive statistical properties that could improve the point-to-grid-cell comparison \citep[see][]{rammig_generic_2018}.f

@@

The initialization of SOC stocks, however, is important for the size of the SOC debt and its change over time, since the presence and size of legacy fluxes affect these values strongly (see Fig. \ref{fig:SOCinit}). According to our sensitivity analysis the SOC debt varies between \(33.3\,\unit{GtC}\) and \(50.7\,\unit{GtC}\) depending on the initialization choice, with our best guess at \(39.6\,\unit{GtC}\). Concurrently, our results indicate that the impact of the dynamic agricultural management is robust to the initialization of SOC stocks.

### 2022-07-20 SOCbudgetPaper/discussion.Rmd
update to final version - 2. resubmit

The Tier 2 approach [@ogle_cropland_in_ipcc_2019] used here is explicitly designed for agricultural soils, whereas we apply it also to soils under PNV. This is necessary in order to represent SOC losses under land-use change in a dynamic way, as this is an important driver of SOC dynamics. The comparison of simulated PNV data with LPJmL4 shows the model's capability in reproducing PNV SOC stocks (Fig. \@ref(fig:SOCLPJmL4)).
Concurrently, the point-data comparison (see Fig. \@ref(fig:SOCpoint)) shows low correlation for PNV, however also for cropland sites. This might also point to the fact that SOC stocks can vary at field and local scale considerably and thus a very high number of point data is needed to derive statistical properties, that could improve the point-to-grid-cell comparison [see @rammig_generic_2018].f

@@

The initialization of SOC stocks, however, is important for the size of the SOC debt and its change over time, since the presence and size of legacy fluxes effect these values strongly (see Fig. \@ref(fig:SOCinit)). According to our sensitivity analysis the SOC debt depending on the initialization choice has been between $33.3\,\unit{GtC}$ and $50.7\,\unit{GtC}$, with our best guess at $39.6\,\unit{GtC}$. Concurrently, our results indicate, that the impact of the dynamic agricultural management is robust to the initialization of SOC stocks.

>>>>

The Tier 2 approach [@ogle_cropland_in_ipcc_2019] used here is explicitly designed for agricultural soils, whereas we apply it also to soils under PNV. This is necessary in order to represent SOC losses under land-use change in a dynamic way, as this is an important driver of SOC dynamics. The comparison of simulated PNV data with LPJmL4 shows the model's capability in reproducing PNV SOC stocks (Fig. \@ref(fig:SOCLPJmL4)).
Concurrently, the point-data comparison (see Fig. \@ref(fig:SOCpoint)) shows low correlation for PNV, however also for cropland sites. This might also point to the fact that SOC stocks can vary at field and local scale considerably and thus a very high number of point data is needed to derive statistical properties that could improve the point-to-grid-cell comparison [see @rammig_generic_2018].f

@@

The initialization of SOC stocks, however, is important for the size of the SOC debt and its change over time, since the presence and size of legacy fluxes affect these values strongly (see Fig. \@ref(fig:SOCinit)). According to our sensitivity analysis the SOC debt varies between $33.3\,\unit{GtC}$ and $50.7\,\unit{GtC}$ depending on the initialization choice, with our best guess at $39.6\,\unit{GtC}$. Concurrently, our results indicate that the impact of the dynamic agricultural management is robust to the initialization of SOC stocks.
